---
title: "Homework 3: covid-19"
subtitle: "STAT 6210 AU Fall 2020"
author: "Ana Gabriela Itokazu"
output:
     prettydoc::html_pretty:
     theme: cayman
highlight: github
---

```{r chunksetup, include=FALSE}
knitr::opts_chunk$set(
   echo=TRUE, eval=TRUE, include=TRUE, warning=FALSE, message=FALSE, fig.align="center"
)
```

```{r clear_genv, echo=FALSE}
rm(list = ls())
```

<div style="text-align: justify">

### The `covid.csv` file contains data on Covid-19 cases in a province of Italy. For the following exercises, only provide first rows or elements in the output if the results are too large.

</div>

```{r install, echo=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
```

<div style="text-align: justify">

I made sure that the `.csv` file was in the same working directory as my project, so there was no file path in the command. I make the first row of the `raw_covid` data frame as the columns names, and the first column, as the rows names. I also re-named the "Epidemiological.link...Notes" column as "Notes", just for the sake of tidiness...

</div>

```{r import}

raw_covid <- read.csv(file = "covid.csv", row.names=1, header=TRUE)

colnames(raw_covid)[8] <- "Notes"

```

```{r view_raw, echo=FALSE}

glimpse(raw_covid)

```
\

<div style="text-align: justify">

### 1 and 2: Correctly format the dates in “Date.of.birth”, “First.day.of.symptoms”, “Date.of.outcome” and “Date.of.diagnosis”; Format “Hospitalization.type”, “Symptoms” and “Outcome” as factors and “Epidemiological.link…Notes” as character.

Because all modifications were in the same object, I made them using piping (`%>%`) and the `mutate` function for each type of modification. For the modifications in more than one variable, I used the `across` function applied to a vector of them.

</div>

```{r q1q2}

raw_covid <- raw_covid %>%
     mutate(across(c(Date.of.birth,
                     First.day.of.symptoms,
                     Date.of.outcome,
                     Date.of.diagnosis), as.Date)) %>%
     mutate(across(c(Hospitalization.type,
                     Symptoms,Outcome), as.factor)) %>%
     mutate(Notes = as.character(Notes))

```

```{r view_q1q2, echo=FALSE}

glimpse(raw_covid)

```
<div style="text-align: justify">
\

### 3: Create a three-dimensional table reporting the three factors from the previous question.

For this, I did not used the `table` function. Instead, I selected the data based on the class of the variable. To use said function, I would do something that looks like the `#commented` line below.
</div>

```{r q3}

q3 <- select_if(raw_covid, is.factor)

# q3 <- table(raw_covid[c("Hospitalization.type",
#                          "Symptoms",
#                          "Outcome")])

```

```{r view_q3, echo=FALSE}

glimpse(q3)

```
\

<div style="text-align: justify">

### 4: Using the table object created in the previous question, subset it in order to create the table presented in the homework assignment.

To be able to reproduce the table, I `filtered` the data to the set represented there. The whole table is only representing the data for the "Asymptomatic", and it does not represent the data that includes "Deceased" or "Non-Intensive care hospitalization". Once filtering all this, I could `group_by` the "Hospitalization Type" and "Outcome". This, summarized by the "Symptoms" column, gives us the counts for this overlap. In a separate piping, I use the function `spread` to move the "Outcome" values in a column based on the "Symptoms" values. Now, this looks like the table provided already, but because the `summarize` function did not provided a zero-value count, we need to `replace` the NA values for zero. This is long, and honestly can be done in like one line of subsetting, but I find it easier to understand and use when I know exactly what I'm taking out and modifying. Also, if I need to use the same function to extract different types of information, it is easier to change small bits. Like that, I can filter out different levels, or group by different variables.

</div>

```{r q4}

 q4 <- q3 %>%
      filter(Symptoms != "Symptomatic") %>%
      filter(Outcome != "Deceased") %>%
      filter(Hospitalization.type != "Non-Intensive care hospitalization") %>%
      group_by(Hospitalization.type, Outcome) %>%
      summarize(Symptoms = n())
 
 q4 <- q4 %>%
      spread(Outcome,Symptoms) %>%
      replace(is.na(.), 0)
 
```

```{r view_q4, echo=FALSE, fig.align="center"}

colnames(q4) <- c("","Disease in progress","Healed")
knitr::kable(q4, align = "c",
             format = "html",
             table.attr = "style='width:75%;'") %>%
        kableExtra::kable_styling(position = "center")

```
\

<div style="text-align: justify">

### 5: Order the data based on the date of diagnosis (from first to most recent).

</div>

```{r q5}

q5 <- arrange(raw_covid, Date.of.diagnosis)

```

```{r view_q5, echo=FALSE}

glimpse(q5$Date.of.diagnosis)

```
\

<div style="text-align: justify">

### 6: Add a column that reports whether or not a case was asymptomatic AND in home isolation. Name the observation “Home_Asymptomatic” if the conditions apply and “Non_Home_Asymptomatic” if not and then produce a bar plot of this new variable.

</div>

```{r q6, echo=TRUE, eval=TRUE, fig.align="center", warning=FALSE, message=FALSE}

df_q6 <- raw_covid %>%
   mutate(q6 = character(nrow(raw_covid)))

for (i in 1:nrow(df_q6)){
   if(isTRUE((df_q6[i,"Hospitalization.type"] == "Home isolation")
   && (df_q6[i, "Symptoms"] == "Asymptomatic"))) {
      df_q6[i,"q6"] <- "Home_Asymptomatic"
      } else {
      df_q6[i,"q6"] <- "Non_Home_Asymptomatic"
      }
}

```

```{r view_q6, fig.align="center"}

plot_q6 <- table(df_q6$q6)

barplot(plot_q6, xlab="Asymptomatic",
        names.arg=c("Home","Non-Home"),
        ylim=c(0,1000))
axis(2,at=seq(0,1000,200))

```
\

<div style="text-align: justify">

### 7: Count the number of cases of people born after 1981 and that have healed.

</div>

```{r q7}

q7 <- raw_covid %>%
     filter(Date.of.birth >= "1981-01-01") %>%
     filter(Outcome == "Healed") 

```

```{r view_q7, echo=FALSE}
q7 <- summarize(q7, n())
colnames(q7) <- "Number of people born after 1981 and healed"
knitr::kable(q7, align = "c",
             format = "html",
             table.attr = "style='width:65%;'") %>%
        kableExtra::kable_styling(position = "center")
```
\

<div style="text-align: justify">

### 8: Count the number of cases that are asymptomatic OR in home isolation (but not both) AND were born before 1982.

</div>

```{r q8, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

q8 <- raw_covid %>%
   filter(Date.of.birth >= "1982-01-01") %>%
   filter(xor(Symptoms == "Asymptomatic",
              Hospitalization.type == "Home isolation"))

```

```{r view_q8, echo=FALSE}
q8 <- summarize(q8, n())
colnames(q8) <- "Number of people born after 1982, Asymptomatic or in Home isolation"
knitr::kable(q8, align = "c",
             format = "html",
             table.attr = "style='width:65%;'") %>%
        kableExtra::kable_styling(position = "center")
```
\

<div style="text-align: justify">

### 9: Create a new dataset including only the rows where “Epidemiological.link…Notes” includes the words “contact” OR “symptom” (or both). Hint: you can use the grep() function and tolower().

Here I used the `tolower` function to make sure I will not mismatch due to differences on typing

</div>

```{r q9, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

q9 <- raw_covid %>%
     filter(grepl("contact", tolower(Notes)) |
            grepl("symptom", tolower(Notes)))

```

<div style="text-align: justify">

### 10: In the previous dataset add a column reporting the age (in years, therefore in integer format) of each patient as of October 2nd, 2020. Save this dataset into a .csv file and make it available on your GitHub repository for this assignment.

</div>

```{r q10, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}


```

<div style="text-align: justify">

### 11: Produce a pie chart for the type of hospitalization for cases born between 1960 and 1980.

</div>

```{r q11, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}


```