---
title: "Homework 3: Salary"
subtitle: "STAT 6210 AU Fall 2020"
author: "Ana Gabriela Itokazu"
output:
     prettydoc::html_pretty:
     theme: cayman
highlight: github
---

```{r chunksetup, include=FALSE}
knitr::opts_chunk$set(
   echo=TRUE, eval=TRUE, include=TRUE,
   warning=FALSE, message=FALSE,
   fig.align="center"
)
```

```{r clear_genv, echo=FALSE}
rm(list = ls())
```

<div style="text-align: justify">

### The `covid.csv` file contains data on Covid-19 cases in a province of Italy. For the following exercises, only provide first rows or elements in the output if the results are too large.

</div>

```{r install, echo=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(tibble)
```

<div style="text-align: justify">

### 1: Load the data and check the formatting of the variables.

I made sure that the `.csv` file was in the same working directory as my project, so there was no file path in the command. I make the first row of the `raw_salary` data frame as the columns names, and the first column, as the rows names.

</div>

```{r q1}

raw_salary <- read.csv(file = "adult.csv",
                       sep = ";",
                       row.names = 1,
                       header = TRUE)

```

```{r view_q1, echo=FALSE}

glimpse(raw_salary)

```
\

<div style="text-align: justify">

### 2: Rename the last column (currently called “NA”) containing the dichotomous salary information. Assign the name salary to it.

</div>

```{r q1q2}

colnames(raw_salary)[15] <- "salary"

```

```{r view_q1q2, echo=FALSE}

glimpse(raw_salary[15])

```
<div style="text-align: justify">
\

### 3: The values in salary have a space in front of them (e.g. " >50K"): remove the space from all values (hint: you can use substring()).

Because it was a leading space, I choose to use `trimws()`. The argument `which` could be `"right"` for trailing spaces, `"both"` for leading and trailing and to remove all spaces.

</div>

```{r q3}

raw_salary$salary <- trimws(raw_salary$salary, which = "left")

```

```{r view_q3, echo=FALSE}

glimpse(raw_salary[15])

```
\

<div style="text-align: justify">

### 4: Again, in the variable salary, replace “>50K” with the value 1 and “<=50K” with 0. Make sure to format it as a factor with two levels.

Here, I chose to factorize the `salary` variable, giving it two levels with the labels "0" and "1".

</div>

```{r q4}

raw_salary$salary <- factor(raw_salary$salary, levels = c("<=50K",">50K"), labels = c(0,1))

```

```{r view_q4, echo=FALSE}

glimpse(raw_salary$salary)

```
\

<div style="text-align: justify">

### 5: Use the glm() function to estimate this logistic model. Only specify the arguments formula = salary ~ ., data and family = binomial. Save the result of the estimation in an object called fit.

</div>

```{r q5}

fit <- glm(salary ~.,data=raw_salary,family=binomial)

```
\

<div style="text-align: justify">

### 6: Using the information in fit (hint: you can use the functions coef() and summary() to extract information), create a dataframe collecting: The names of the variables (names of the rows of the dataframe); The value of the coefficients (first column); A logical vector stating which coefficients are positive (second column); The p-values (third column).

Again, I did it in a long way, and I'm sure there are ways of doing it in less lines. This piece of code extracts the information using `summary.glm`, which outputs a matrix with the values of `Estimates`, `Standard Errors`, `z-value`, and `p-values` for all variables. In the first line, I proceeded to extract only the `Estimates` and `p-values` columns. The second line binds this matrix with a new column that states a logical vector for values of `Estimates` bigger than zero (positive). The third line simply renames the TRUE/FALSE for Positive/Negative for tidiness. Finally, the last line reorders the data frame to the requested order.

</div>

```{r q6}

fit_results <- as.data.frame(summary.glm(fit)$coefficients[,c(1,4)])

q6 <- cbind(fit_results, coeff=fit_results[,1] > 0)

q6$coeff <- factor(q6$coeff,
                   levels = c("FALSE","TRUE"),
                   labels = c("Negative","Positive"))

q6 <- q6[c(1,3,2)]

```

```{r view_q6, echo=FALSE}

colnames(q6) = c("Coefficients","q6","p-values")

kable(head(q6))

```
\

<div style="text-align: justify">

### 7: Subset the dataframe created in the previous question to only show the rows where the p-values are strictly smaller than 0.05. Knowing that (i) the remaining rows are statistically significant and that (ii) positive coefficients contribute to increase the probability of a salary larger than $50,000 (the opposite for negative values), comment on those variables that negatively contribute to salary..

Code-wise, I simply filter the `p-values` smaller than 0.05.

</div>

```{r q7}

q7 <- q6 %>%
        filter(`p-values` < 0.05)

```

```{r view_q7, echo=FALSE}

q7 %>%
        kable(booktabs = T) %>%
        kable_styling() %>%
        row_spec(which(q7$q6 == "Negative"),
                 color = "black", background = "white") %>%
        row_spec(which(q7$q6 == "Positive"),
                 color = "black", background = "lightgrey") 

```


